---
title: "Assignment 3"
author: "Suraj Gangaram"
date: "2/25/2022"
output: html_document
---

```{r}
source("analysis.R")
```

# Introduction

Through analyzing the incarceration data from the past five decades, I have computed
five summary statistics. For the most part, I am focusing on the incarceration
levels regarding the black population with respect to different contexts (year, 
state, etc.).The **first** one is the state with the largest amount of incarcerations
in the most recent year of the dataset, the year 2018. The **second** statistic 
refers to which individual year the black jail population was the highest in the 
country. The **third** one explores the ratio of the black population in Georgia 
during 2018 to that of the black jail population in Georgia in 2018. The **fourth** 
one is the ratio of the black jail population in 2018 to that of the total population nationwide in 2018. The **fifth** one dives into the race which had been the most incarcerated in 2018.

## Questions

**Q1**: Which _state_ had the largest amount of black citizens in jail in 2018?
**A1**: The state and population (state, population) is `r state_highest_black_total`.

**Q2**: In which _year_ was the black jail population the highest in US?
**A2**: The year and population (year, population) `r year_highest_black_total`.

**Q3**: What is the _ratio_ of the black population (from ages 15-64) in Georgia 
in 2018 to the black jail population in Georgia in 2018?
**A3**: The ratio is `r black_ratio_2018`.

**Q4**: Which was the _most_ incarcerated _race_ in 2018?
**A4**: The race and population (race, population) is `r race_most_incarcerated_data`.

**Q5**: What is the _ratio_ of the black jail population nationwide in 2018 to
the total population nationwide in 2018?
**A5**: The ratio is `r black_country_ratio_2018`.

## Plots: Line Chart

```{r line_chart, echo=FALSE}
plot(line_chart)
```

I included this chart in order to visualize the differences in incarceration among
all of the major races denoted in the dataset. I decided to put the jail population
as the y-variable and the year as the x-variable to take note of the incarceration
population trends for each race throughout the years. I was surprised to find that 
blacks, which typically is the race associated with high incarceration _rates_, had relatively low jail populations.

## Plots: Scatterplot

```{r scatter_chart, echo=FALSE}
plot(scatter_chart)
```

I included this plot in order to visualize one of the variables in the 3rd question
(black jail population of Georgia) so the viewer can understand the trend of the 
black jail population throughout the years. It seems like that the dataset only 
included data on the black jail population of Georgia from 1985 onwards, and the 
population peaked in around 2009, the year when President Obama was inaugurated
(which makes me wonder if there is a correlation between the two as the amount of
black people in Georgia jails dwindles from there).

## Plots: Map

```{r black_map, echo=FALSE}
plot(black_map)
```

I included this map in order to visualize the black jail population in 2018 (which
is part of the 5th question) so the viewer can understand how many black people
were incarcerated in each state. The states with high black populations (California,
Texas, Georgia, etc.) have the highest black jail population. The converse seems
true with states such as Vermont and Delaware with a small black population having 
little to no black jail population.

